---
title: 'Sporulation assay with IPTG-induced sigma factors '
output: github_document
editor_options: 
  chunk_output_type: inline
  code_folding: hide
---

```{r setup, , include=FALSE}
library(renv)
library(here)
library(tidyverse, quietly = TRUE)
library(cowplot, quietly = TRUE)
library(outliers, quietly = TRUE)
library(broom)
library(gtools)
```


```{r load data, echo=FALSE, message=FALSE}
# get all the file names to collect
files <- list.files(here("data/output"), pattern = ".csv",
                    full.names = T, recursive = T)
d <- tibble()

for (f in files){
   # tmp <- read_csv(f)
   # if (!"clust.model" %in% colnames(tmp)){
   #    tmp$clust.model <- "general"
   # }
   d <- rbind(d, read_csv(f))
   
      }

# # confirm layout
# d %>%
#    # filter(exp == "exp4") %>%
#    mutate(r = str_extract(well,"^."), c=parse_number(well)) %>%
#    ggplot(aes(c,r))+
#    geom_tile(aes(fill = colony))+
#    geom_text(aes(label =rep))+
#    facet_wrap(.~exp)+
#    scale_y_discrete(limits=rev)+
#    scale_x_continuous(breaks = 1:12)+
#    scale_fill_viridis_d()



d <- d%>%
   mutate(strain=str_replace(d$strain,"Eldridge","ELDg"))%>%
   mutate(strain = fct_relevel(strain, "ELDg168", "ELDg169","pDR110","sigF", "sigG","SP10","Goe3"))

d$treat <- recode(d$treat, "xIPTG"="noIPTG")

```

**Only consider counts with at least 100 events**

Quantities based on lower event counts are designated a value of 1 cell/mL, to prevent problems with ratio and logs.
```{r clean data, echo=FALSE}
d <- d %>% 
   mutate(spore.ml = if_else(spore < 100, 1, spore.ml)) %>% 
   mutate(veg.ml = if_else(veg < 100, 1, veg.ml))
```


# Replication outliers
I have N=3 for each flask at each time point. There are some measurements which are obviously way off. To get rid of these I will choose from each triplicate the 2 points that are in best agreement and remove the 3rd point. There is a function to idenify such a point in pacakge 'outliers'
> outlier {outliers} Finds value with largest difference between it and sample mean, which can be an outlier.

> logical: if set to TRUE, gives vector of logical values, and possible outlier position is marked by TRUE

I will apply the outlier filtering on the number of total cells (veg+spore)

```{r id outliers, echo=FALSE}
ol <- 
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   mutate(cells.ml = spore.ml+veg.ml) %>% 
   select(well,rep.group,cells.ml)
ol$outlier <- NA 
for (i in unique(ol$rep.group)){

   ol$outlier[ol$rep.group==i]=
   ol%>%
      filter(rep.group==i)%>%
      select(cells.ml)%>%
      deframe()%>%
      outlier(logical = TRUE)
  
}

```

# Overview of results

Concentrations of cell types: 
```{r plot overview, message=F, echo=FALSE}
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","SP10","Goe3","ELDg168", "ELDg169"))%>%
   pivot_longer(names_to = "pop", values_to = "cell.ml", cols = c(veg.ml, spore.ml))%>%
   group_by(strain, treat,colony,pop,exp)%>%
   summarise(cell.ml=mean(cell.ml))%>%
   ungroup()%>%
   mutate(exp=as_factor(exp))%>%
   
   ggplot(aes(x=strain, y=cell.ml))+
      geom_col(aes(group=interaction(colony,exp), fill=exp), position = position_dodge(width = 0.8), color="black")+
      stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.8), size=1.5, width=0.5, color="blue")+
   stat_summary(geom = "point", fun = mean, position = position_dodge(width = 0.8), shape=21, fill="blue", size=3) +
   
   facet_grid(pop~treat)+
   scale_y_log10()+
   annotation_logticks(side="l", color="grey")+
   ylab("Number per mL (log)")+
   theme_classic(base_size = 14)+
   panel_border(color = "black")+
   labs(caption = "mean±SE")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   scale_fill_viridis_d()+
   coord_cartesian(ylim = c(1e5, NA))

```
Remove blanks
```{r Remove blanks, echo=FALSE}
d <- d %>% 
   filter(strain != "blank")

```


## Total cells

```{r sum total cells, echo=FALSE}
d.sum.cells <- d%>%
   #remove outlier for technical replicates
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","Goe3","ELDg168","SP10", "ELDg169"))%>%
   mutate(cells.ml=veg.ml+spore.ml)%>%
   group_by(strain, treat,colony,exp)%>%
   summarise(cells.ml=mean(cells.ml) ,
             spore.ml=mean(spore.ml), 
             veg.ml = mean(veg.ml) )%>%
   pivot_longer(cols = ends_with("ml"), names_to = "pop", values_to = "quant") %>%  
     ungroup()

d.diff.cells <- d%>%
   #remove outlier for technical replicates
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   
   select("strain","exp", "colony","treat","spore.ml","veg.ml")%>%
   mutate(cells.ml=veg.ml+spore.ml)%>%
      pivot_longer(cols = ends_with("ml"), names_to = "pop", values_to = "quant") %>% 
   group_by(strain, treat,colony,exp, pop)%>%
   summarise(quant = mean(quant) )%>%
   pivot_wider(names_from = "treat", values_from = "quant") %>% 
  ungroup() %>% 
     mutate(diff = noIPTG - IPTG)

```


```{r stats total cells}
# summary(aov(induction.ratio~strain+colony+exp, d.sum.cells))
   

   d.test.cells <- tibble()
# t-tests against control
   for (s in unique(d.sum.cells$strain)){
      for (who in c("spore.ml", "veg.ml", "cells.ml")){
              
          d.test.cells <- d.sum.cells %>% 
               filter(strain == s, pop == who) %>% 
               t.test(quant ~ treat, data = .) %>% 
               broom::tidy() %>% 
               mutate(strain = s, pop = who) %>% 
               bind_rows(d.test.cells, .)
      }

   }
   
# adjust p-value for multiple testing
   d.test.cells <- d.test.cells %>% 
      mutate(adj.p = p.adjust(p.value, method = "BH"),
             p.lab = stars.pval(adj.p)) %>% 
      relocate(strain, p.value, adj.p, p.lab)
   
   knitr::kable(d.test.cells, format = "html")
```


```{r stats diff cells}

   d.test.diffs <- tibble()
# t-tests against control
   for (s in unique(d.diff.cells$strain)){
      for (who in c("spore.ml", "veg.ml", "cells.ml")){
              
          d.test.diffs <- d.diff.cells %>% 
               filter(strain == s, pop == who) %>% 
             pull(diff) %>% 
               t.test(., paired = F, mu = 0) %>% 
               broom::tidy() %>% 
               mutate(strain = s, pop = who) %>% 
               bind_rows(d.test.diffs, .)
      }

   }
   
# adjust p-value for multiple testing
   d.test.diffs <- d.test.diffs %>% 
      mutate(adj.p = p.adjust(p.value, method = "BH"),
             p.lab = stars.pval(adj.p)) %>% 
      relocate(strain, p.value, adj.p, p.lab)
   
   knitr::kable(d.test.diffs, format = "html")
```

```{r cell stats for plot}
plot.p <- 
d.test.cells%>% 
   select(strain, pop, p.value, p.lab) %>% 
   
   mutate(strain = str_replace(strain, "pDR110", "empty vector")) %>% 
   mutate(strain = fct_relevel(strain,"empty vector","sigF", "sigG","SP10", "Goe3","ELDg168", "ELDg169"))%>%
   
      mutate(pop = if_else(pop=="spore.ml", "spores","vegetative")) %>%
    mutate(pop = fct_relevel(pop,"spores", "vegetative")) %>% 
 
   
      # panel separation
   mutate(pnl=case_when(strain == "empty vector" ~ "neg. ctrl",
                        strain %in% c("sigF","sigG") ~ "host",
                        TRUE ~ "phage") %>% 
             as_factor()) %>% 
   mutate(pnl = fct_relevel(pnl,"phage", "host", "neg. ctrl")) %>%
   mutate(y = if_else(pop=="vegetative", -9e7, 9e7)) %>% 
   mutate(p.lab = str_replace(p.lab, "\\.", ""))

```


```{r plot total cells, echo=FALSE}

d.plot.cells <- 
d%>%
   #remove outlier for replicates
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   #reduce table
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = str_replace(strain, "pDR110", "empty vector")) %>% 
   mutate(strain = fct_relevel(strain,"empty vector","sigF", "sigG","SP10", "Goe3","ELDg168", "ELDg169"))%>%

   # average technical replicates
   group_by(strain, treat,colony,exp)%>%
   summarise(spore.ml=mean(spore.ml), veg.ml = mean (veg.ml)) %>% 
   
   # long format
   pivot_longer(cols = c("spore.ml", "veg.ml"),
                names_to = "pop", values_to = "cell.ml") %>% 
   mutate(pop = if_else(pop=="spore.ml", "spores","vegetative")) %>%
    mutate(pop = fct_relevel(pop,"spores", "vegetative")) %>% 
 
   
      # panel separation
   mutate(pnl=case_when(strain == "empty vector" ~ "neg. ctrl",
                        strain %in% c("sigF","sigG") ~ "host",
                        TRUE ~ "phage") %>% 
             as_factor()) %>% 
   mutate(pnl = fct_relevel(pnl,"phage", "host", "neg. ctrl")) %>%
   
   mutate(`IPTG induction` = ifelse(treat=="noIPTG", "-", "+")) %>% 
   
   mutate(cell.ml = if_else(pop=="vegetative", -cell.ml, cell.ml))



axis.breaks <- c(-1e8, -5e7, 0, 5e7, 1e8)
   
p <- 
   d.plot.cells %>% 
   # average biological replicates
   group_by(pnl, strain, `IPTG induction`, pop)%>%
   summarise(m = mean(cell.ml), n = n(), v = sd(cell.ml)/sqrt(n),
             .groups = "drop")%>%

   #plot
   ggplot(aes(x=`IPTG induction`, y=m, fill = pop))+
   geom_hline(yintercept = 0, color="grey")+
   geom_bar(position="stack", stat="identity", width = 0.5, color="black")+
   geom_errorbar(position="identity", width = 0.2 , size = 1,
                 aes(ymin = m-v, ymax = m+v))+
   geom_text(data = plot.p, aes(x = 1.5, y = y, label = p.lab, color = pop),
             size=8, alpha = 0.5)+
   # geom_jitter(data = d.plot.cells, aes(x=`IPTG induction`, y = cell.ml),
   #            alpha = 0.5,show.legend = FALSE, shape=21, width = 0.2, fill="white")+

   facet_wrap(pnl~strain, nrow = 2)+

   ylab("cells/ml")+
   scale_y_continuous(breaks = axis.breaks, labels = abs(axis.breaks),
                      limits = c(-1.1e8,1.1e8))+
   scale_fill_viridis_d(direction = 1)+
   scale_colour_viridis_d(direction = 1)+
   theme_classic(base_size = 14)+
   panel_border(color="black")+
   labs(caption = "mean±SE (n ≥ 8) ; P-values: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘ ’ 1")+
   theme(axis.text.x = element_text(size = 18),
         legend.position = "bottom")

ggsave(here("fig/induction_response_cells.png"),p,
       width = 6, height = 6)
p


```


```{r plot diff cells, echo=FALSE}

d.plot.diff <- 
d.diff.cells%>%
   select(-diff) %>% 
   filter(strain != "blank") %>% 
   pivot_longer(cols = contains("IPTG"), names_to = "treat", values_to = "cell.ml") %>% 

   # panel separation
   mutate(strain = str_replace(strain, "pDR110","empty vector" )) %>% 
   mutate(pnl=case_when(strain == "empty vector" ~ "neg. ctrl",
                        strain %in% c("sigF","sigG") ~ "host",
                        TRUE ~ "phage") %>% 
             as_factor()) %>% 
   mutate(pnl = fct_relevel(pnl,"phage", "host", "neg. ctrl")) %>%
   
   mutate(`IPTG induction` = ifelse(treat=="noIPTG", "-", "+")) 



p <-
   d.plot.diff %>% 
   # average biological replicates
   group_by(pnl, strain, `IPTG induction`, pop)%>%
   summarise(m = mean(cell.ml), n = n(), v = sd(cell.ml)/sqrt(n),
             .groups = "drop")%>%

   #plot
   ggplot(aes(x=`IPTG induction`, y=m, color = pop))+
   geom_line(data = d.plot.diff, aes(y=cell.ml, group = interaction(exp,colony,pop)), alpha = 0.2, size = 0.5)+
   geom_line(aes(group = pop))+

   geom_line(aes(group = pop), size =1)+
   geom_pointrange(aes(ymin = m-v, ymax = m+v))+#shape = 21,fill = "white")+

   facet_wrap(pnl~strain, nrow = 2)+

   ylab("cells/ml")+
  scale_y_log10()+
   scale_fill_viridis_d(direction = 1)+
   scale_colour_viridis_d(direction = 1)+
   theme_classic(base_size = 14)+
   panel_border(color="black")+
   labs(caption = "mean±SE (n ≥ 8)" )+
   theme(axis.text.x = element_text(size = 18),
         legend.position = "bottom")

ggsave(here("fig/induction_response.png"),p,
       width = 6, height = 6)
p


```

Change in response to induction, log10(IPTG/noIPTG):

```{r plot response, message=F, echo=FALSE}
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","SP10","Goe3","ELDg168", "ELDg169"))%>%
   pivot_longer(names_to = "pop", values_to = "cell.ml", cols = c(veg.ml, spore.ml))%>%
   group_by(strain, treat,colony,pop,exp)%>%
   summarise(cell.ml=mean(cell.ml))%>%
   pivot_wider(names_from = treat, values_from = cell.ml)%>%
   mutate(induction.ratio.log=log10(IPTG/noIPTG))%>%
   ungroup()%>%
   mutate(exp=as_factor(exp))%>%
   ggplot(aes(x=strain, y=induction.ratio.log))+
      geom_col(aes(group=interaction(colony,exp), fill=exp), position = position_dodge(width = 0.8), color="black")+
      stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.8), size=1.5, width=0.5, color="blue")+
   stat_summary(geom = "point", fun = mean, position = position_dodge(width = 0.8), shape=21, fill="blue", size=3) +
   facet_grid(.~pop)+geom_hline(yintercept = 0)+
   ylab("log ( IPTG/noIPTG )")+
   theme_classic(base_size = 14)+
   panel_border(color = "black")+
   labs(caption = "mean±SE")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   scale_fill_viridis_d()+
   coord_cartesian(ylim = c(NA, 2))

```


```{r, echo=FALSE}
p <- 
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
    mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","Goe3","ELDg168","SP10", "ELDg169"))%>%
   pivot_longer(names_to = "pop", values_to = "cell.ml", cols = c(veg.ml, spore.ml))%>%
   group_by(strain, treat,colony,pop,exp)%>%
   summarise(cell.ml=mean(cell.ml))%>%
   pivot_wider(names_from = treat, values_from = cell.ml)%>%
   mutate(induction.ratio.log=log10(IPTG/noIPTG))%>%
   ungroup()%>%
   mutate(exp=as_factor(exp))%>%
   mutate(pop=fct_rev(pop))%>%
   ggplot(aes(x=strain, y=induction.ratio.log))+
      stat_summary(geom = "col", fun = mean, position = position_dodge(width = 0.8), fill="grey", color="black") +
         stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.8), size=1, width=0.2)+
         geom_point(aes(group=interaction(colony,exp), shape=exp), ,position = position_jitterdodge(jitter.width = NULL, jitter.height = 0,
  dodge.width = 0.3), size=2)+
   # stat_summary(geom = "point", fun = mean, position = position_dodge(width = 0.8), shape=21, fill="white", size=3) +
   facet_grid(.~pop)+geom_hline(yintercept = 0)+
   ylab("log ( IPTG/noIPTG )")+
   theme_classic(base_size = 14)+
   panel_border()+
   ggtitle("Response to induction (mean±SE)")+
   scale_shape_manual(values = c(21:25))+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "bottom")+
   annotation_logticks(side="l", color="grey")

# ggsave("induction_response.pdf",p)

p
```

```{r, echo=FALSE}
p <- 
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","Goe3","ELDg168","SP10", "ELDg169"))%>%
   mutate(perc.spore=spore.ml/(veg.ml+spore.ml))%>%
   group_by(strain, treat,colony,exp)%>%
   summarise(perc.spore=mean(perc.spore))%>%
   pivot_wider(names_from = treat, values_from = perc.spore)%>%
   mutate(induction.ratio.log=log10(IPTG/noIPTG))%>%
   ungroup()%>%
   mutate(exp=as_factor(exp))%>%
   # mutate(pop=fct_rev(pop))%>%
   
   ggplot(aes(x=strain, y=induction.ratio.log))+
      stat_summary(geom = "col", fun = mean, position = position_dodge(width = 0.8), fill="grey", color="black") +
         stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.8), size=1, width=0.2)+
         geom_point(aes(group=interaction(colony,exp), shape=exp), ,position = position_jitterdodge(jitter.width = NULL, jitter.height = 0,
  dodge.width = 0.3), size=2)+
   # stat_summary(geom = "point", fun = mean, position = position_dodge(width = 0.8), shape=21, fill="white", size=3) +
   # facet_grid(.~pop)+geom_hline(yintercept = 0)+
   ylab("log ( IPTG/noIPTG )")+
   theme_classic(base_size = 14)+
   panel_border()+
   ggtitle("Response to induction %spores (mean±SE)")+
   scale_shape_manual(values = c(21:25))+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "bottom")+
   annotation_logticks(side="l", color="grey")

# # ggsave("induction_response_perc.pdf",p)
# library (officer)
# library(rvg)
# read_pptx() %>%
#   add_slide(layout = "Two Content", master = "Office Theme") %>%
#   ph_with(dml(ggobj = p), location = ph_location_type(type = "body")) %>%
#   print(target = "plot.pptx")
p

```

#stats on log ratio of % spores

```{r sum log-ratio, echo=FALSE}
d.sum <- d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","Goe3","ELDg168","SP10", "ELDg169"))%>%
   mutate(perc.spore=spore.ml/(veg.ml+spore.ml))%>%
   group_by(strain, treat,colony,exp)%>%
   summarise(perc.spore=mean(perc.spore))%>%
   pivot_wider(names_from = treat, values_from = perc.spore)%>%
   mutate(induction.ratio=(IPTG/noIPTG))%>%
   ungroup()
```


```{r stats log-ratio}
summary(aov(induction.ratio~strain+colony+exp, d.sum))
   

   d.test <- tibble()
# t-tests against control
   for (s in unique(d.sum$strain)){
      if (s =="pDR110") next
      d.test <-d.sum %>% 
         filter(strain == s |strain =="pDR110") %>% 
         t.test(induction.ratio ~ strain, data = .) %>% 
         broom::tidy() %>% 
         mutate(strain = s) %>% 
         bind_rows(d.test, .)
   }
   
# adjust p-value for multiple testing
   d.test <- d.test %>% 
      mutate(adj.p = p.adjust(p.value, method = "BH"),
             p.lab = stars.pval(adj.p)) %>% 
      relocate(strain, p.value, adj.p, p.lab)
   
   knitr::kable(d.test, format = "html")
```


```{r plot log-ratio, echo=FALSE}
p <- 
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","Goe3","ELDg168","SP10", "ELDg169"))%>%
   mutate(perc.spore=100*spore.ml/(veg.ml+spore.ml))%>%
   group_by(strain, treat,colony,exp)%>%
   summarise(perc.spore=mean(perc.spore))%>%
   # pivot_wider(names_from = treat, values_from = perc.spore)%>%
   ungroup()%>%
   mutate(exp=as_factor(exp))%>%
   # mutate(pop=fct_rev(pop))%>%
   
   ggplot(aes(x=strain, y=perc.spore))+
      stat_summary(geom = "col", fun = mean, position = position_dodge(width = 0.8), fill="grey", color="black") +
         stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.8), size=1, width=0.2)+
         geom_point(aes(group=interaction(colony,exp), shape=exp), position = position_jitterdodge(jitter.width = NULL, jitter.height = 0,
  dodge.width = 0.3), size=2)+
   # stat_summary(geom = "point", fun = mean, position = position_dodge(width = 0.8), shape=21, fill="white", size=3) +
   facet_grid(.~treat)+geom_hline(yintercept = 0)+
   ylab("%spores")+
   theme_classic(base_size = 14)+
   panel_border()+
   # ggtitle("Response to induction %spores (mean±SE)")+
   scale_shape_manual(values = c(21:25))+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "bottom")+
   ylim(0,100)

# ggsave("perc_spores.pdf",p)

p
```


# ms plot

```{r, echo=FALSE}
d.plot <- 
d%>%
   #remove outlier for replicates
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   #reduce table
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
   mutate(strain = str_replace(strain, "pDR110", "empty vector")) %>% 
   mutate(strain = fct_relevel(strain,"empty vector","sigF", "sigG","SP10", "Goe3","ELDg168", "ELDg169"))%>%
   # calc percent spore
   mutate(perc.spore=spore.ml/(veg.ml+spore.ml))%>%
   
   # average technical replicates
   group_by(strain, treat,colony,exp)%>%
   summarise(perc.spore=mean(perc.spore)) %>% 
   
      # panel separation
   mutate(pnl=case_when(strain == "empty vector" ~ "neg. ctrl",
                        strain %in% c("sigF","sigG") ~ "host",
                        TRUE ~ "phage") %>% 
             as_factor()) %>% 
   mutate(pnl = fct_relevel(pnl,"phage", "host", "neg. ctrl"))




   
d.plot %>% 
   # average biological replicates
   group_by(pnl, strain, treat)%>%
   summarise(m = mean(perc.spore), n = n(), v = sd(perc.spore)/sqrt(n),
             .groups = "drop")%>%
   
   #plot
   ggplot(aes(x=strain, y=m, fill = fct_rev(treat)))+
   # geom_col( position = position_dodge(width = 0.5), 
   #          width = 0.5, color="black")+
      geom_rect(aes(ymin = 5e-3, ymax = m,
                 xmin = ..x..-.3, xmax = ..x..+.3),
                   position = position_dodge(width = 0.5), width = 0.5,# fill = fct_rev(treat)),
                color= "black", show.legend = F)+
   geom_errorbar(aes(ymin = m-v, ymax = m+v, group = fct_rev(treat)),
                 position = position_dodge(width = 0.5), width = 0.5)+
   geom_point(data = d.plot, aes(x=strain, y = perc.spore),alpha = 0.5,
              position = position_jitterdodge(dodge.width =  0.5, jitter.width = 0.05),
              show.legend = FALSE, shape=21)+
   
   facet_grid(.~pnl, scales = "free_x", space = "free_x")+

   ylab("perc. spores")+
   scale_y_continuous(labels = scales::percent, trans = "log10")+
   scale_fill_viridis_d(direction = 1)+
   theme_classic(base_size = 14)+
   panel_border(color="black")+
   labs(caption = "mean±SE")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "bottom")

# # ggsave("induction_response_perc.pdf",p)
# library (officer)
# library(rvg)
# read_pptx() %>%
#   add_slide(layout = "Two Content", master = "Office Theme") %>%
#   ph_with(dml(ggobj = p), location = ph_location_type(type = "body")) %>%
#   print(target = "plot.pptx")
# p

```

```{r plot log col, echo=FALSE}
d.test.plot <- 
   d.test %>% 
   mutate(p.lab = str_replace(p.lab, "\\.", "")) %>% 
   # panel separation
   mutate(pnl=case_when(#strain == "empty vector" ~ "neg. ctrl",
                        strain %in% c("sigF","sigG") ~ "host",
                        TRUE ~ "phage") %>% 
             as_factor()) %>% 
   mutate(pnl = fct_relevel(pnl,"phage", "host"))

d.plot.log <- d.plot %>% 
   pivot_wider(names_from = treat, values_from = perc.spore)%>%
   mutate(induction.ratio=(IPTG/noIPTG))%>%
   ungroup()

p <- d.plot.log %>% 
   # average biological replicates
   group_by(pnl, strain)%>%
   summarise(m = mean(induction.ratio), n = n(),
             v = sd(induction.ratio)/sqrt(n), .groups = "drop")%>%
   
   #plot
   ggplot(aes(x=strain, y=m) )+
   # geom_col( position = position_dodge(width = 0.5), 
   #          width = 0.5, color="black")+
   geom_rect(aes(ymin = 1e-5, ymax = m,
                 xmin = ..x..-.3, xmax = ..x..+.3,
                 fill=pnl), color= "black", show.legend = F)+
   geom_errorbar(aes(ymin = m-v, ymax = m+v), width = 0.3)+
   geom_point(data = d.plot.log, aes(x=strain, y = induction.ratio, shape = exp),
              alpha = 0.7,position = position_jitter(width = 0.15),
              show.legend = FALSE, fill = "white")+
   #p values asterisks
   geom_text(data = d.test.plot, aes( label = p.lab, y = 1.2),
             size = 10, color = "grey") + 
   
   facet_grid(.~pnl, scales = "free_x", space = "free_x")+

   ylab("spore yield")+
   xlab("induced sigma factor")+
   scale_y_continuous(labels = scales::percent, trans = "log10")+
   annotation_logticks(sides = "l", colour = "grey")+
   coord_cartesian(ylim = c(5e-3,NA))+
   scale_fill_viridis_d()+
   scale_shape_manual(values = 21:25)+
   theme_classic(base_size = 12)+
   panel_border(color="black")+
   labs(caption = "mean±SE ; P-values compared with empty vector: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘ ’ 1")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "bottom",
         plot.caption = element_text(colour = "grey"))

ggsave(here("fig/induction_response_perc.png"),p,
       width = 6, height = 4)
p

```

```{r, echo=FALSE}
p <- 
d%>%
   mutate(rep.group=interaction(strain, treat,colony,exp))%>%
   left_join(., ol%>%select(well, outlier, rep.group), by=c("well", "rep.group"))%>%
   filter(!outlier)%>%
   select("strain","treat", "colony","rep","spore.ml","veg.ml","exp")%>%
    mutate(strain = fct_relevel(strain,"pDR110","sigF", "sigG","SP10", "Goe3","ELDg168", "ELDg169"))%>%
   mutate(perc.spore=spore.ml/(veg.ml+spore.ml))%>%
   group_by(strain, treat,colony,exp)%>%
   summarise(perc.spore=mean(perc.spore))%>%
   pivot_wider(names_from = treat, values_from = perc.spore)%>%
   mutate(induction.ratio.log=(IPTG/noIPTG))%>%
   ungroup()%>%
   mutate(exp=as_factor(exp))%>%
        # panel separation
   mutate(pnl=case_when(strain == "pDR110" ~ "neg.ctrl",
                        strain %in% c("sigF","sigG") ~ "host",
                        TRUE ~ "phage") %>% 
             as_factor()) %>% 
   mutate(pnl = fct_relevel(pnl,"phage", "host", "neg.ctrl")) %>% 
   
   ggplot(aes(x=strain, y=induction.ratio.log))+
   geom_hline(yintercept = 1)+
      stat_summary(geom = "col", fun = mean, position = position_dodge(width = 0.8), fill="grey", color="black", width = 0.5) +
         stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.8), size=1, width=0.3)+
         geom_point(aes(group=interaction(colony,exp), shape=exp), position = position_jitterdodge(jitter.width = NULL, jitter.height = 0,
  dodge.width = 0.2), size=2, show.legend = F)+
   
   facet_grid(.~pnl, scales = "free_x", space = "free_x")+
   ylab("spore yield")+
   xlab("sigma factor gene expressed")+
   scale_y_log10()+
   theme_classic(base_size = 14)+
   panel_border(color="black")+
   labs(caption = "mean±SE")+
   scale_shape_manual(values = c(21:25))+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "bottom")
# # ggsave("induction_response_perc.pdf",p)
# library (officer)
# library(rvg)
# read_pptx() %>%
#   add_slide(layout = "Two Content", master = "Office Theme") %>%
#   ph_with(dml(ggobj = p), location = ph_location_type(type = "body")) %>%
#   print(target = "plot.pptx")
p
```